Linear Mixed Models

### 2023-04-22 3_HowTos_LinearMixedModels.qmd
added link to pdf version of manuscript

# ex1.interceptonly {.unnumbered}

<style>
div.blue pre { background-color:lightblue; }
div.blue pre.r { background-color:lightgrey; }

.nav-tabs > li.active:nth-child(1) > a {
  background-color: ##964B00;
}

.nav-tabs > li > a {
  background-color: ##FFFF00;
  color: blue;
}

main-container {
   max-width: unset;
}
.descriptives {
   overflow: auto; /*do not use `resize`*/
}

.tableby {
   overflow: auto;
}

.descriptives {
   overflow: auto;
   resize: both;
}

table, th, td {
  padding-top: 0px !important;
  padding-bottom: 0px !important;
  vertical-align: middle;
}
</style>

::: column-body-outset
## Linear Mixed Model with a Random-Intercept (LMM:ri)
<p>
The **`LMM:ri`** is similar to traditional (fixed-effect) linear regression extending on this approach by including a subject-specific random-effect that allows each participant to have their own unique intercept value, in addition to the overall mean-level (fixed-effect) intercept value

In this example, we will use the **`LMM:ri`** to analyze trajectories of scores on the externalizing subscale of the child behavior checklist (CBCL) obtained across three measurement occasions in a sample of youth taking part in the ABCD Study. Our primary aim is to characterize stability and change in CBCL externalizing scores across assessments, while accounting for observations that are clustered within youth over time. To do so, we will use the **`LMM:ri`** to simultaneously model an overall sample mean trajectory (fixed effect) and subject-specific (random) effects that vary randomly about the sample mean trajectory. 
</p>

:::
:::

###############################

::: {.column-body-outset}
## Preliminary Setup
::: {.panel-tabset}
### Load Libraries {.tabset .tabset-fade .tabset-pills}
example-text before code block
<div class = "blue">
```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220), fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
</div>
:::
:::

::: {.column-body-outset}
## Descriptives Overview
::: {.panel-tabset}
### Data Table {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r data-table}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

```
</div>
:::
:::

::: {.column-body-outset}
## Results 
::: {.panel-tabset}
### Build Model {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r build-model}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

#df_long %>%
#  report() %>%
#  summary()
```
</div>
:::
:::

::: {.column-body-outset}
## Visualizations 
::: {.panel-tabset}
### Extract Params {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r extrct-FE}
#| messages: FALSE
#| warning: FALSE

@@

```
</div>
:::
:::

@@

::: column-body-outset
#### Interpretation
The estimated correlation between x and y was "r cor(x,y)`.
There are "r nrow(my_data)` individuals.

>>>>

## Linear Mixed Model with a Random-Intercept (LMM:ri)
<p>
The LMM:ri is similar to traditional (fixed-effect) linear regression extending on this approach by including a subject-specific random-effect that allows each participant to have their own unique intercept value, in addition to the overall mean-level (fixed-effect) intercept value zzzzt

In this example, we will use the LMM:ri to analyze trajectories of scores on the externalizing subscale of the child behavior checklist (CBCL) obtained across three measurement occasions in a sample of youth taking part in the ABCD Study. Our primary aim is to characterize stability and change in CBCL externalizing scores across assessments, while accounting for observations that are clustered within youth over time. To do so, we will use the LMM:ri to simultaneously model an overall sample mean trajectory (fixed effect) and subject-specific (random) effects that vary randomly about the sample mean trajectory. 
</p>

## Preliminary Setup
::: {.panel-tabset}
### Load Libraries {.tabset .tabset-fade .tabset-pills}
example-text before code block
<div class = "blue">
```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220), fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
</div>
:::

## Descriptives Overview
::: {.panel-tabset}
### Data Table {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r data-table}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

```
</div>
:::

## Results 
::: {.panel-tabset}
### Build Model {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r build-model}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

#df_long %>%
#  report() %>%
#  summary()
```
</div>
:::

## Visualizations 
::: {.panel-tabset}
### Extract Params {.tabset .tabset-fade .tabset-pills}
<div class = "blue">
```{r extrct-FE}
#| messages: FALSE
#| warning: FALSE

@@

```
</div>
:::

@@

#### Interpretation
The estimated correlation between x and y was "r cor(x,y)`.
There are "r nrow(my_data)` individuals.
